Simulated Observations of Young Protoplanetary Discs

### 2013-01-24 paper.tex
figs edited, text slightly changed

\author[T.~A.~Douglas et al.]
{\parbox{\textwidth}{Tom Douglas$^{1}$\thanks{E-mail: \texttt{pytd@leeds.ac.uk}},
Paola Caselli$^{1}$,
Aaron Boley$^{2}$,
Richard Durisen$^{3}$,
Tom Hartquist$^{1}$,
John Ilee$^{1}$ and
Jonathon Rawlings$^{4}$\\
\vspace{0.1cm}\\
{\small{\it$^{1}$School of Physics and Astronomy, University of Leeds, Leeds LS2 9JT, UK}} \\
{\small{\it$^{2}$Department of Astronomy, University of Florida, 211 Bryant Space Center, PO Box 112055, USA}}\\
{\small{\it$^{3}$Department of Astronomy, Indiana University, 727 East 3rd Street, Swain West 319, Bloomington, IN 47405, USA}}\\
{\small{\it$^{4}$Department of Physics \& Astronomy, University College London, London WC1E 6BT, UK}}\\}}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.eps}
 \label{grid}
 \caption{A plot of the points selected by the griding process and the paths down which photons can propagate overlaid on a smoothed density model. The points are more concentrated at small radii and in the densest regions.}
\end{figure}

@@

Simulations are limited to molecules for which we have abundances in both the disc and envelope, and also have calculated Einstein and collisional coefficients. The Molecular data used is from the lamda database (reference? http://home.strw.leidenuniv.nl/$\sim$moldata/). Lines for the molecules which satisfy these conditions were simulated with LIME for frequencies between 50$\sim$500$\,$GHz, with a focus on lines which could be observed in ALMA band 7 and molecules which are not typically seen in outflows. The results presented in this section represent those lines which show detectable emission/absorption which can be used to trace either spiral structure or rotation.\newline

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.eps}
 \label{points}
 \caption{A 300$\,$GHz continuum image of the model, created from the average of ten LIME runs.}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageOCS_28-27_30deg_composite_all.eps}
 \label{OCS_all}
 \caption{OCS 28-27 {\bf Left:} Continuum subtracted integrated intensity map showing clearly differentiated spiral structure in the inner regions of the disc, {\bf Right:} Intensity weighted velocity map}
\end{figure*}

@@

\begin{figure*}
  \includegraphics[width=168mm]{Figures/sim/imageHCOp_1-0_30deg_composite_all.eps}  \caption{HCO$^+$ 1-0 {\bf Left:} Continuum subtracted integrated intensity map, {\bf Right:} Intensity weighted velocity map}
  \label{hco+_all}
\end{figure*}

@@

In order to make predictions on the observability of the synthetic brightness maps created by LIME, CASA (common astronomy software applications) was used to simulate observing these with the completed ALMA in configuration 26 of 28 for a total time of six hours, giving a beam size of between 0.02 and 0.03 arcseconds. In order to clean the images which featured mainly absorption the sky model used was the output of the lime simulations with the continuum subtracted and then the image inverted. The simulated images were cleaned down to a 3$\sigma$level with $\sigma$given by the on-line ALMA sensitivity calculator (ref).

Sensitivities:\newline
C17O J=3$\rightarrow$2 at 337.06GHz $\sigma$=0.0270$\,$mJy$\,$beam$^{-1}$ \newline
OCS J=26$\rightarrow$25 at 316.15GHz $\sigma$=0.0275$\,$mJy$\,$beam$^{-1}$ \newline
HCO$^+$ J=3$\rightarrow$2 at 267.56GHz $\sigma$=0.0176$\,$mJy$\,$beam$^{-1}$ \newline
H$_2$CO J=4$_{04}\rightarrow$3$_{03}$ at 290.62GHz $\sigma$=0.0228$\,$mJy$\,$beam$^{-1}$ \newline

@@

Figure \ref{h2co_chanmap} shows a channel map of the H$_2$CO 4$_{04}$-$3_{03}$ line overlaid on the 1$\,$mm continuum emission. From this it can be seen that the absorption features trace the disc spiral structure whilst line emission emanates from gaps in the disc where there is less continuum to be absorbed.{\bf not sure what else to say about channel map ask paola}

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/casa_all_30deg_contSub.eps}
 \caption{Integrated intensity maps for the simulated observations of:{\bf Top Left:} H$_2$CO 4$_{04}\rightarrow$3$_{03}$, {\bf Top Right:}C$^{17}$O 3$\rightarrow$2, {\bf Bottom Left:} OCS 26$\rightarrow$25 and {\bf Bottom Right:} HCO$^+$ 3$\rightarrow$2. contours start at 3$\sigma$ and increasing in intervals of 3$\sigma$ for each of the maps except the H$_2$CO which starts at 5$\sigma$ and increases in intervals of 5$\sigma$.}
\label{mom0_maps}
\end{figure*}

>>>>

\author[T.~A.~Douglas et al.]
{\parbox{\textwidth}{Tom Douglas$^{1}$\thanks{E-mail: \texttt{pytd@leeds.ac.uk}},
Paola Caselli$^{1}$,
Aaron Boley$^{2}$,
Richard Durisen$^{3}$,
Tom Hartquist$^{1}$,
John Ilee$^{1}$ and
Jonathan Rawlings$^{4}$\\
\vspace{0.1cm}\\
{\small{\it$^{1}$School of Physics and Astronomy, University of Leeds, Leeds LS2 9JT, UK}} \\
{\small{\it$^{2}$Department of Astronomy, University of Florida, 211 Bryant Space Center, PO Box 112055, USA}}\\
{\small{\it$^{3}$Department of Astronomy, Indiana University, 727 East 3rd Street, Swain West 319, Bloomington, IN 47405, USA}}\\
{\small{\it$^{4}$Department of Physics \& Astronomy, University College London, London WC1E 6BT, UK}}\\}}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/grid.eps}%Lime_grid3.eps
 \label{grid}
 \caption{A plot of the points selected by the griding process and the paths down which photons can propagate overlaid on a smoothed density model. The points are more concentrated at small radii and in the densest regions.}
\end{figure}

@@

Simulations are limited to molecules for which we have abundances in both the disc and envelope, and also have calculated Einstein and collisional coefficients. The Molecular data used is from the lamda database (Sch\"oier et al. (2005) http://home.strw.leidenuniv.nl/$\sim$moldata/). Lines for the molecules which satisfy these conditions were simulated with LIME for frequencies between 50$\sim$500$\,$GHz, with a focus on lines which could be observed in ALMA band 7 and molecules which are not typically seen in outflows. The results presented in this section represent those lines which show detectable emission/absorption which can be used to trace either spiral structure or rotation.\newline

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.eps}
 \caption{{\bf Left:}A 300$\,$GHz continuum image of the model, created from the average of ten LIME runs. {\bf Right:} the output of a single radiative transfer simulation, with artefacts due to finite gridding.}
 \label{averages}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageOCS_26-25_30deg_all.eps}
 \caption{OCS 28-27 {\bf Left:} Continuum subtracted integrated intensity map showing clearly differentiated spiral structure in the inner regions of the disc, {\bf Right:} Intensity weighted velocity map}
 \label{OCS_all}
\end{figure*}

@@

\begin{figure*}
  \includegraphics[width=168mm]{Figures/sim/imageHCOp_3-2_30deg_all.eps}  \caption{HCO$^+$ 3-2 {\bf Left:} Continuum subtracted integrated intensity map, {\bf Right:} Intensity weighted velocity map}
  \label{hco+_all}
\end{figure*}

@@

In order to make predictions on the observability of the synthetic brightness maps CASA (common astronomy software applications) was used to simulate their observtion with the completed ALMA in configuration 26 of 28, giving a beam size of between 0.02 and 0.03 arcseconds. As the weakest lines studied in the model are of the order of 0.1-0.2$\,$mJy$\,$beam$^{-1}$ at this resolution, sensitivities of the order of 0.02$\,$mJy$\,$beam$^{-1}$ are required. This requires around 6 hours of integration time, the exact sensitivities are give in table \ref{sigmas}.

\begin{table}
 \centering
 \begin{minipage}{80mm}
   \caption{Sensitivities of ALMA given 6 hours integration time. As calculated by the ALMA on-line sensitivity calculator.}
   \label{sigmas}
   \begin{tabular}{c||c|c|c}
     \hline
     Species & Transition & Frequency & Sensitivity\\
     \hline
     C$^{17}$O & J=3$\rightarrow$2 & 337.06GHz & 0.0270$\,$mJy$\,$beam$^{-1}$ \\
     OCS & J=26$\rightarrow$25 & 316.15GHz & 0.0275$\,$mJy$\,$beam$^{-1}$ \\
     HCO$^+$ & J=3$\rightarrow$2 & 267.56GHz & 0.0176$\,$mJy$\,$beam$^{-1}$ \\
     H$_2$CO & 4$_{04}\rightarrow$3$_{03}$ & 290.62GHz & 0.0228$\,$mJy$\,$beam$^{-1}$ \\
     \hline
   \end{tabular}
 \end{minipage}
\end{table}

 In order to clean the images which featured mainly absorption the sky model used was the output of the lime simulations with the continuum subtracted and then the image inverted. The simulated images were cleaned down to a 3$\sigma$level with $\sigma$ given in table \ref{sigmas}.

@@

Figure \ref{h2co_chanmap} shows a channel map of the H$_2$CO 4$_{04}$-$3_{03}$ line overlaid on the 1$\,$mm continuum emission. From this it can be seen that the absorption features trace the disc spiral structure whilst line emission emanates from gaps in the disc where there is less continuum to be absorbed.

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/casa_all_30deg_contSub.eps}
 \caption{Integrated intensity maps for the simulated observations of:{\bf Top Left:} H$_2$CO 4$_{04}\rightarrow$3$_{03}$, {\bf Top Right:}C$^{17}$O 3$\rightarrow$2, {\bf Bottom Left:} OCS 26$\rightarrow$25 and {\bf Bottom Right:} HCO$^+$ 3$\rightarrow$2. contours start at 3$\sigma$ and increasing in intervals of 3$\sigma$ for each of the maps except the H$_2$CO which starts at 5$\sigma$ and increases in intervals of 5$\sigma$. Emission is in red contours, absorption in blue.}
\label{mom0_maps}
\end{figure*}
